---
title: "p8105_hw6_kmw2189"
author: "Kylie Wheelock Riley"
date: "11/25/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## Problem 1
```{r rerading and viewing prob 1, output = FALSE}
## load in data
birthweight_df = read_csv("data/birthweight.csv") %>% 
  janitor::clean_names()

## checking for missings, etc.
skimr::skim(birthweight_df)
##there appear to be no missings for any of the the variables, yay!
```


```{r cleaning and tidying prob 1}
##changing the var type for babysex, mrace, and frace
birthweight_df = birthweight_df %>% 
  mutate(
  sex = as.factor(babysex),
  sex = recode(sex, "1" = "male", "2" = "female" ),
  m_race = as.character(mrace),
  m_race = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
  f_race = as.character(frace),
  f_race = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown")
  ) %>% 
  select(-babysex, -mrace, -frace) %>% 
  select(sex, everything()) 
 
```

**Model fitting**

From my experience working with birth outcomes, specifically birthweight, there are 3 covariates that are absolutely necessary to include in the model: gestational age, maternal BMI, and a socioeconomic status indicator. For this model I will use family income as the SES indicator. 
```{r}
birthwt_mdl_1 = lm(bwt ~ gaweeks + ppbmi + fincome, data = birthweight_df) 

birthwt_mdl_1 %>% 
  broom::tidy() 

```

```{r}
birthweight_df %>% 
  modelr::add_residuals(birthwt_mdl_1) %>% 
  modelr::add_predictions(birthwt_mdl_1) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()

```

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
```{r}
birthwt_mdl_2 = lm(bwt ~ gaweeks + blength, data = birthweight_df) 

birthwt_mdl_2 %>% 
  broom::tidy() 
```

One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
birthwt_mdl_3 = lm(bwt ~ bhead + blength + sex + bhead*blength + bhead*sex + blength*sex + bhead*blength*sex, data = birthweight_df) 

birthwt_mdl_3 %>% 
  broom::tidy() 
```

