---
title: "p8105_hw6_kmw2189"
author: "Kylie Wheelock Riley"
date: "11/25/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(10)

```

## Problem 1

**Loading and tidying data**

```{r reading and reviewing data, results = 'hide'}
## load in data
birthweight_df = read_csv("data/birthweight.csv") %>% 
  janitor::clean_names()

## checking for missings, etc.
skimr::skim(birthweight_df)
##there appear to be no missings for any of the the variables, yay!
```


```{r cleaning and tidying prob 1}
##changing the var type for babysex, mrace, and frace
birthweight_df = birthweight_df %>% 
  mutate(
  sex = as.factor(babysex),
  sex = recode(sex, "1" = "male", "2" = "female" ),
  m_race = as.character(mrace),
  m_race = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
  f_race = as.character(frace),
  f_race = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown")
  ) %>% 
  select(-babysex, -mrace, -frace) %>% 
  select(sex, everything()) 
 
```

**Model fitting**

From my experience working with birth outcomes, specifically birthweight, there are 3 covariates that are necessary to include in the model: gestational age, maternal BMI, and a socioeconomic status (SES) indicator. For this model I will use family income as the SES indicator. 

```{r fitting model 1}
birthwt_mdl_1 = lm(bwt ~ gaweeks + ppbmi + fincome, data = birthweight_df)
```

```{r visualization of model 1}
##plot of model residuals against fitted values
birthweight_df %>% 
  modelr::add_residuals(birthwt_mdl_1) %>% 
  modelr::add_predictions(birthwt_mdl_1) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point(alpha = 0.5) +
  labs(
      title = "Model 1 residuals against predicted values",
      x = "Predicted Values",
      y = "Residuals",
      caption = "Summary of predicted values from a model of birthweight by the residuals of the model"
  )
```

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
```{r creating model 2}
birthwt_mdl_2 = lm(bwt ~ gaweeks + blength, data = birthweight_df) 

```

One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r creating model 3}
birthwt_mdl_3 = lm(bwt ~ bhead + blength + sex + bhead*blength + bhead*sex + blength*sex + bhead*blength*sex, data = birthweight_df) 

```

**Comparing Models**

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r plotting models}
##Plotting predicted birthweight values against the true birthwight values for each model
birthweight_df %>% 
  gather_predictions(birthwt_mdl_1, birthwt_mdl_2, birthwt_mdl_3) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = bwt, y = pred))  + 
  geom_point() + 
  ##geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```


```{r using modelr for cross validation}
##create the training and testing datasets using modelr
cv_df =
  crossv_mc(birthweight_df, 100)

##use mutate + map & map2 to fit models to training data and obtain corresponding RMSEs for the testing data
cv_df = 
  cv_df %>% 
  mutate(
    birthwt_mdl_1 = map(train, ~lm(bwt ~ gaweeks + ppbmi + fincome, data = .x)), 
    birthwt_mdl_2 = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
    birthwt_mdl_3 = map(train, ~lm(bwt ~ bhead + blength + sex + bhead*blength + bhead*sex + blength*sex + bhead*blength*sex, data = as_tibble(.x)))
    ) %>% 
  mutate(
    rmse_mdl_1 = map2_dbl(birthwt_mdl_1, test, ~rmse(model = .x, data = .y)),
    rmse_mdl_2 = map2_dbl(birthwt_mdl_2, test, ~rmse(model = .x, data = .y)),
    rmse_mdl_3 = map2_dbl(birthwt_mdl_3, test, ~rmse(model = .x, data = .y))
    )
```

```{r plotting RMSEs}
##plotting rmse for each model
cv_df %>% 
  select(starts_with("rmse")) %>% 
pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(
      title = "RMSEs for each model",
      x = "Model",
      y = "RMSE",
      caption = "Evaluating the best fit of models predicting birthweight"
  )
```

**Conclusion**

My initially predicted model using gestational age, maternal BMI, and income as an SES measure did not perform nearly as well as I anticipated! Model 3, or the "fully loaded" model containing head circumference, birth length and sex interactions performed the best. If I were to revise my original model I would definitely include birth length. 

## Question 2

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1). 

**Load data and look at plot**

```{r loading prob 2 data, warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

##graph data
weather_df %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm") 
```

**Bootstrap using modelr**

```{r}
bootstrap_df = 
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)), 
    rsquared = map(models, broom::glance),
    results = map(models, broom::tidy)
  ) %>% 
  select(-strap, -models) %>% 
  unnest(rsquared, results) %>% 
  pivot_wider(
    id_cols = c(.id, r.squared),
    names_from = "term", 
    values_from = "estimate"
    ) %>% 
  janitor::clean_names() %>% 
  rename(
    beta_0 = intercept, 
    beta_1 = tmin 
  ) %>% 
  mutate(
    beta_log = log(beta_0 * beta_1)
  )
  
```

Plot the distribution of your estimates
```{r}
bootstrap_df %>%
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  labs(
      title = "Distribution of R Squared",
      x = "R Squared Values",
      y = "Density"
  )

bootstrap_df %>% 
  ggplot(aes(x = beta_log)) + 
  geom_density() +
    labs(
      title = "Distribution of Log (Beta 0 * Beta 1)",
      x = "Log (Beta 0 * Beta 1)",
      y = "Density"
  )
  
```

Both the distribution of the r squared values and the log of Beta 0  x Beta 1 appear to be normal.  A linear regressionwould be appropriate to use with the data. 


Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1)

```{r}
summary_vars_df = 
  bootstrap_df %>%
  mutate(
    r_squared = as.vector(r_squared),
    beta_log = as.vector(beta_log)
  ) %>% 
  summarize(
    r_squared_low = quantile(r_squared, probs = (0.025), na.rm=TRUE), 
    r_squared_high = quantile(r_squared, probs = (0.975), na.rm=TRUE), 
    log_beta_low = quantile(beta_log, probs = (0.025), na.rm=TRUE),
    log_beta_high = quantile(beta_log, probs = (0.975), na.rm=TRUE)
  )
  
```

The 95% confidence interval for r squared is `r round(pull(summary_vars_df, r_squared_low), digits = 2)` to `r round(pull(summary_vars_df, r_squared_high), digits = 2)` and the 95% confidence interval for log(β^0∗β^1) is `r round(pull(summary_vars_df, log_beta_low), digits = 2)` to `r round(pull(summary_vars_df, log_beta_high), digits = 2)`




